Carbonate chemistry at Ny-Ålesund: AWIPEV-CO~2~ Project

### 2020-02-21 awipev-CO2.Rmd
temperature insitu + seafet calibration OK

@@

Within this deployment period, calibration is made with E0 calculation (first plot) and final corrected pH is calculated (second plot). To allow calibration, one spectrophotometric pH measurement must match: seaFET voltage, in situ temperature (SBE38) and salinity. When one of these parameters is missing for any reasons (Ferrybox maintenance, breakdown...), the gap is filled with a mean of the 4 previous or next hours when available. If the pH spectrophotometric measurement is too far from the available values, the calibration is not performed (environemental variability reasons). The SBE38 insitu temperature and seaFET temperature are very similar (slope = 0.997, r2 = 0.948). These two temperatures were used interchangeably if one is missing.

@@

# We decided to calibrate pH data within each pH spectro measurement.
# Thus, each pH spectro is taken as one calibration periode. 
# We calculate here the day between 2 pH spectro in order to have the start/end of one deployment period.
sfPeriod <- d_all%>%
  dplyr::select(datetime, pH_s_seafet_temp_field)%>%
  dplyr::filter(!is.na(pH_s_seafet_temp_field))

@@

#put interval in a list with the start date of seaFET measurements
sfinterv <- c( as_datetime("2017-08-24 13:00:00"), sfPeriod[,4])
 #Change period sfinterv[6]. Because seafet was off and on on the 2018-04-24 19:00:00
sfinterv[6] <- ymd_hms("2018-04-24 18:00:00")
 #Change period sfinterv[6]. Because seafet was off and on on the 2018-04-24 19:00:00
sfinterv[14] <- ymd_hms("2019-01-29 16:00:00")

@@

# # 5th period, with pH spectro at 2018-04-16 07:00:00 must fit one voltINT to calculate E0
# we put the mean of 4 following days because voltINT was not present.
mean_voltINT_4d <- d_all%>%
  dplyr::filter(datetime >= "2018-04-25" & datetime <= "2018-04-29")%>%
  dplyr::summarise( mean(voltINT, na.rm = TRUE))
mean_voltINT_4d <- as.numeric(mean_voltINT_4d)
d_all <- d_all%>%
  dplyr::mutate(voltINT = ifelse(datetime == "2018-04-16 07:00:00" , mean_voltINT_4d, voltINT))

mean_voltEXT_4d <- d_all%>%
  dplyr::filter(datetime >= "2018-04-25" & datetime <= "2018-04-29")%>%
  dplyr::summarise( mean(voltEXT, na.rm = TRUE))
mean_voltEXT_4d <- as.numeric(mean_voltEXT_4d)
d_all$voltEXT <- ifelse(d_all$datetime == "2018-04-16 07:00:00" , mean_voltEXT_4d, d_all$voltEXT)

# put also a salinity value because FB was off
sal_to_replace <- d_all%>%
  dplyr::filter(datetime >= "2018-04-16 00:00:00" & datetime <= "2018-04-16 23:59:59")%>%
  dplyr::summarise( mean(sal_fb, na.rm = TRUE))
sal_to_replace <- as.numeric(sal_to_replace)
d_all$sal_fb <- ifelse(d_all$datetime == "2018-04-16 07:00:00" , sal_to_replace, d_all$sal_fb)
#interpolation of salinity during the FB off maintenance
sal_fbinterp <-approx(d_all$datetime, d_all$sal_fb, xout=d_all$datetime, method="linear", rule=2)
names(sal_fbinterp) <- c("datetime", "sal_fb_interp")
sal_fbinterp <- as.data.frame(sal_fbinterp)
d_all$sal_fb_interp <- sal_fbinterp$sal_fb_interp

@@

# put also a temp value because FB was off
temp_to_replace <- d_all%>%
  dplyr::filter(datetime >= "2018-04-16 00:00:00" & datetime <= "2018-04-16 23:59:59")%>%
  dplyr::summarise( mean(temp_fb, na.rm = TRUE))
temp_to_replace <- as.numeric(temp_to_replace)
d_all$temp_fb <- ifelse(d_all$datetime == "2018-04-16 07:00:00" , temp_to_replace, d_all$temp_fb)

# put also a temp insi value because FB was off FOR PERIOD 1 : 2017-08-24
temp_to_replace <- d_all%>%
  dplyr::filter(datetime >= "2017-08-24 12:00:00" & datetime <= "2017-08-24 23:00:00")%>%
  dplyr::summarise( mean(T_seaF, na.rm = TRUE))
temp_to_replace <- as.numeric(temp_to_replace)
d_all$temp_insitu <- ifelse(d_all$datetime == "2017-08-24 14:00:00" , temp_to_replace, d_all$temp_insitu)

test <- d_all%>%
  dplyr::filter(datetime=="2017-08-24 14:00:00")

@@

#@@@@@ test plot #@@@@@ #@@@@@ #@@@@@ 
seafet_contros_xts <- dplyr::select(d_all, datetime, voltINT, phINT_filtered,pH_s_seafet_temp_field, T_seaF, temp_fb, temp_insitu, sal_fb)
#seafet_contros_xts <- dplyr::select(d_all, datetime,voltINT)
seafet_contros_xts <- as.xts(seafet_contros_xts, order.by = d_all$datetime)
dygraph(seafet_contros_xts, group = "awipev", main="pH seaFET (in situ)", ylab="pHT") %>%
  dySeries("voltINT", label = "voltINTINT", color ="blue", strokeWidth=0, pointSize=2) %>%
    dySeries("pH_s_seafet_temp_field", label = "pH spectro", axis= "y2", color =  "green", strokeWidth = 0, pointSize = 4 ) %>%
    dySeries("phINT_filtered", label = "Raw pH INT", color ="red",axis= "y2", strokeWidth=0, pointSize=2) %>%
  dySeries("temp_fb", label = "temp_fb", color ="orange",axis= "y2", strokeWidth=0, pointSize=2) %>%

  dySeries("sal_fb", label = "sal_fb", color ="pink",axis= "y2", strokeWidth=0, pointSize=2) %>%
  dySeries("temp_insitu", label = "temp_insitu", color ="black",axis= "y2", strokeWidth=0, pointSize=2) %>%
    dySeries("T_seaF", label = "T_seaF", color ="grey",axis= "y2", strokeWidth=0, pointSize=2) %>%
  dyLegend(show = "auto")   %>%
  dyAxis( "y",valueRange = c(-1, 2))%>%
  dyHighlight(highlightCircleSize = 8,highlightSeriesBackgroundAlpha = 0.5,hideOnMouseOut = TRUE) %>%
  dyOptions(drawGrid = TRUE, drawPoints = TRUE, pointSize = c(1,8,1),useDataTimezone = TRUE, digitsAfterDecimal=4) %>%
  dyRangeSelector(height = 30, dateWindow= NULL)
#@@@@@ #@@@@@ #@@@@@ #@@@@@ #@@@@@

# Offset between seafet pH and spec pH 
sfoffset <- left_join(sfPeriod[,c(1,2)], d_all%>%dplyr::select(datetime, pHint_tot_sf), by="datetime")
dat <- mutate(sfoffset, diff_pH_sf = pH_s_seafet_temp_field - pHint_tot_sf)
average_offset <- mean(abs(dat$diff_pH_sf), na.rm=TRUE)

@@

### **T in situ / T seaFET**
```{r Temp comparaison, echo=FALSE, message= FALSE, warning=FALSE}
fit <- lmodel2(data = d_all,  T_seaF ~ temp_insitu , nperm = 99)

pTseaf_Tinsi<-  ggplot(d_all,aes(x=temp_insitu, y= T_seaF, label = month))+
  geom_text( aes(color=factor(month))) + 
 # scale_color_viridis_d(option = "C", guide="none")+
  geom_line(aes(x = T_seaF, y = T_seaF), linetype = "dashed") +
  geom_abline(aes(intercept = fit$regression.results[2,2], slope = fit$regression.results[2,3]),colour = "blue") +
  labs(title = paste("T seaF vs T insi","\n Adj R2 = ", signif(fit$rsquare, 3),
                     "; Intercept =", signif(fit$regression.results[2,2], 3),
                     "; Slope =", signif( fit$regression.results[2,3], 3),
                     "; P =", signif(fit$P.param, 3))) +
  theme(aspect.ratio=1,plot.title = element_text(size=7)) + 
  coord_fixed(ratio = 1 )+ scale_colour_discrete(guide="none") +
  labs( x="T insi", y="T seaF ", color = "Months")  +theme_bw()

print(pTseaf_Tinsi)
```

>>>>

## Temperatures
Several temperature sensors are on site, 1 inside the Ferrybox and 2 outside:

- 1 SBE45 inside the Ferrybox (temp_fb)
- 1 seaFET T°C inside the seaFET at 11 meters depth (T_seaF)
- 1 SBE38 at 11 meters depth (temp_insitu)

Comparative approch allowed by following plots:

### **T seaFET / T in situ **
```{r Temp comparaison, echo=FALSE, message= FALSE, warning=FALSE}
# seaF outliers (T - volt int - volt ext) to remove from 2017-12-10 to 2017-12-14
d_all <- d_all %>%
  dplyr::mutate(T_seaF = ifelse(datetime >= "2017-12-10 11:00:00" & datetime <= "2017-12-14 12:00:00", NA, T_seaF),
                voltINT = ifelse(datetime >= "2017-12-10 11:00:00" & datetime <= "2017-12-14 12:00:00", NA, voltINT), 
                voltEXT = ifelse(datetime >= "2017-12-10 11:00:00" & datetime <= "2017-12-14 12:00:00", NA, voltEXT))

fit <- lmodel2(data = d_all,  T_seaF ~ temp_insitu , nperm = 99)

pTseaf_Tinsi<-  ggplot(d_all,aes(x=temp_insitu, y= T_seaF, label = month))+
  geom_text( aes(color=factor(month))) + 
 # scale_color_viridis_d(option = "C", guide="none")+
  geom_line(aes(x = T_seaF, y = T_seaF), linetype = "dashed") +
  geom_abline(aes(intercept = fit$regression.results[2,2], slope = fit$regression.results[2,3]),colour = "blue") +
  labs(title = paste("T seaF vs T insi","\n Adj R2 = ", signif(fit$rsquare, 3),
                     "; Intercept =", signif(fit$regression.results[2,2], 3),
                     "; Slope =", signif( fit$regression.results[2,3], 3),
                     "; P =", signif(fit$P.param, 3))) +
  theme(aspect.ratio=1,plot.title = element_text(size=7)) + 
  coord_fixed(ratio = 1 )+ scale_colour_discrete(guide="none") +
  labs( x="T insi", y="T seaF ", color = "Months")  +theme_bw()
print(pTseaf_Tinsi)

d_all$diff_T_seaF_insitu <- d_all$T_seaF - d_all$temp_insitu
mean_diff_T_seaF_insitu <- round(mean(d_all$diff_T_seaF_insitu, na.rm = TRUE), digits = 2)
sd_diff_T_seaF_insitu <- round(sd(d_all$diff_T_seaF_insitu, na.rm = TRUE), digits = 2)

The mean difference between T seaFET and in situ is: `r mean_diff_T_seaF_insitu` ± `r sd_diff_T_seaF_insitu°C.

### **T Ferrybox / T in situ **
```{r Temp fb insi comparaison, echo=FALSE, message= FALSE, warning=FALSE}
# Remove temp_fb outliers 
d_all <- d_all %>%
  dplyr::mutate(temp_fb = ifelse(datetime >= "2017-03-22 08:00:00" & datetime <= "2017-03-23 16:00:00", NA, temp_fb),
                temp_fb = ifelse(temp_fb >15, NA, temp_fb))

fit <- lmodel2(data = d_all,  temp_fb ~ temp_insitu , nperm = 99)

pTfb_Tinsi<-  ggplot(d_all,aes(x=temp_insitu, y= temp_fb, label = month))+
  geom_text( aes(color=factor(month))) + 
 # scale_color_viridis_d(option = "C", guide="none")+
  geom_line(aes(x = temp_fb, y = temp_fb), linetype = "dashed") +
  geom_abline(aes(intercept = fit$regression.results[2,2], slope = fit$regression.results[2,3]),colour = "blue") +
  labs(title = paste("T FB vs T insi","\n Adj R2 = ", signif(fit$rsquare, 3),
                     "; Intercept =", signif(fit$regression.results[2,2], 3),
                     "; Slope =", signif( fit$regression.results[2,3], 3),
                     "; P =", signif(fit$P.param, 3))) +
  theme(aspect.ratio=1,plot.title = element_text(size=7)) + 
  coord_fixed(ratio = 1 )+ scale_colour_discrete(guide="none") +
  labs( x="T insi", y="T FB ", color = "Months")  +theme_bw()
print(pTfb_Tinsi)

d_all$diff_T_fb_insitu <- d_all$T_seaF - d_all$temp_insitu
mean_diff_T_fb_insitu <- round(mean(d_all$diff_T_fb_insitu, na.rm = TRUE), digits = 2)
sd_diff_T_fb_insitu <- round(sd(d_all$diff_T_fb_insitu, na.rm = TRUE), digits = 2)
```

The mean difference between T seaFET and in situ is: `r mean_diff_T_fb_insitu` ± `r sd_diff_T_fb_insitu`°C.

@@

Within this deployment period, calibration is made with E0 calculation (first plot) and final corrected pH is calculated (second plot). To allow calibration, one spectrophotometric pH measurement must match: seaFET voltage, in situ temperature (SBE38) and salinity. When one of these parameters is missing for any reasons (Ferrybox maintenance, breakdown...), the gap is filled with a mean of the 4 previous or next hours when available. If the pH spectrophotometric measurement is too far from the available values, the calibration is not performed (environemental variability reasons). The SBE38 insitu temperature and seaFET temperature are very similar (slope = 0.992, r2 = 0.96, see tab "Temperatures"). These two temperatures were used interchangeably if one is missing.

@@

# We decided to calibrate pH data within each pH spectro measurement.
# Thus, each pH spectro is taken as one calibration periode. 
# We calculate here the day between 2 pH spectro in order to have the start/end of one deployment period.
sfPeriod <- d_all%>%
  dplyr::select(datetime, pH_s_seafet_temp_field)%>%
  dplyr::filter(!is.na(pH_s_seafet_temp_field))
# Remove pH spectro that not allow a calibration because of missing calibration value (voltINT, Sal or Temp insi)
sfPeriod <- sfPeriod%>%
dplyr::filter(datetime != "2018-04-16 07:00:00" & datetime != "2018-12-14 09:00:00" & datetime != "2019-01-18 09:00:00")

@@

#put interval in a list with the start date of seaFET measurements
sfinterv <- c( as_datetime("2017-08-24 13:00:00"), sfPeriod[,4])
#  #Change period sfinterv[6]. Because seafet was off and on on the 2018-04-24 19:00:00
# sfinterv[6] <- ymd_hms("2018-04-24 18:00:00")
#  #Change period sfinterv[6]. Because seafet was off and on on the 2018-04-24 19:00:00
# sfinterv[14] <- ymd_hms("2019-01-29 16:00:00")

@@

# period 1, mean of 4 next hours for volt int + T seaF
mean_voltINT_4h <- d_all%>%
  dplyr::filter(datetime >= "2017-08-24 19:00:00" & datetime <= "2017-08-24 22:00:00")%>%
  dplyr::summarise( mean(voltINT, na.rm = TRUE))
mean_voltINT_4h <- as.numeric(mean_voltINT_4h)
d_all <- d_all%>%
  dplyr::mutate(voltINT = ifelse(datetime == "2017-08-24 14:00:00" , mean_voltINT_4h, voltINT))

mean_voltEXT_4h <- d_all%>%
  dplyr::filter(datetime >= "2017-08-24 19:00:00" & datetime <= "2017-08-24 22:00:00")%>%
  dplyr::summarise( mean(voltEXT, na.rm = TRUE))
mean_voltEXT_4h <- as.numeric(mean_voltEXT_4h)
d_all <- d_all%>%
  dplyr::mutate(voltEXT = ifelse(datetime == "2017-08-24 14:00:00" , mean_voltEXT_4h, voltEXT))

mean_TseaF_4h <- d_all%>%
  dplyr::filter(datetime >= "2017-08-24 19:00:00" & datetime <= "2017-08-24 22:00:00")%>%
  dplyr::summarise( mean(T_seaF, na.rm = TRUE))
mean_TseaF_4h <- as.numeric(mean_TseaF_4h)
d_all <- d_all%>%
  dplyr::mutate(temp_insitu = ifelse(datetime == "2017-08-24 14:00:00" , mean_TseaF_4h, temp_insitu))

# period 2, mean of 4 next hours for volt int + T seaF
mean_voltINT_4h <- d_all%>%
  dplyr::filter(datetime >= "2017-09-28 15:00:00" & datetime <= "2017-09-28 18:00:00")%>%
  dplyr::summarise( mean(voltINT, na.rm = TRUE))
mean_voltINT_4h <- as.numeric(mean_voltINT_4h)
d_all <- d_all%>%
  dplyr::mutate(voltINT = ifelse(datetime == "2017-09-28 13:00:00" , mean_voltINT_4h, voltINT))

mean_voltEXT_4h <- d_all%>%
  dplyr::filter(datetime >= "2017-09-28 15:00:00" & datetime <= "2017-09-28 18:00:00")%>%
  dplyr::summarise( mean(voltEXT, na.rm = TRUE))
mean_voltEXT_4h <- as.numeric(mean_voltEXT_4h)
d_all <- d_all%>%
  dplyr::mutate(voltEXT = ifelse(datetime == "2017-09-28 13:00:00" , mean_voltEXT_4h, voltEXT))

# period 20, mean of 4 previous hours for volt int + T seaF
mean_voltINT_4h <- d_all%>%
  dplyr::filter(datetime >= "2019-09-13 04:00:00" & datetime <= "2019-09-13 07:00:00")%>%
  dplyr::summarise( mean(voltINT, na.rm = TRUE))
mean_voltINT_4h <- as.numeric(mean_voltINT_4h)
d_all <- d_all%>%
  dplyr::mutate(voltINT = ifelse(datetime == "2019-09-13 08:00:00" , mean_voltINT_4h, voltINT))

mean_voltEXT_4h <- d_all%>%
  dplyr::filter(datetime >= "2019-09-13 04:00:00" & datetime <= "2019-09-13 07:00:00")%>%
  dplyr::summarise( mean(voltEXT, na.rm = TRUE))
mean_voltEXT_4h <- as.numeric(mean_voltEXT_4h)
d_all <- d_all%>%
  dplyr::mutate(voltEXT = ifelse(datetime == "2019-09-13 08:00:00" , mean_voltEXT_4h, voltEXT))

# period 21, mean of 4 previous hours for volt int + T seaF
mean_voltINT_4h <- d_all%>%
  dplyr::filter(datetime >= "2019-09-13 11:00:00" & datetime <= "2019-09-13 12:00:00")%>%
  dplyr::summarise( mean(voltINT, na.rm = TRUE))
mean_voltINT_4h <- as.numeric(mean_voltINT_4h)
d_all <- d_all%>%
  dplyr::mutate(voltINT = ifelse(datetime == "2019-09-13 13:00:00" , mean_voltINT_4h, voltINT))

mean_voltEXT_4h <- d_all%>%
  dplyr::filter(datetime >= "2019-09-13 11:00:00" & datetime <= "2019-09-13 12:00:00")%>%
  dplyr::summarise( mean(voltINT, na.rm = TRUE))
mean_voltEXT_4h <- as.numeric(mean_voltEXT_4h)
d_all <- d_all%>%
  dplyr::mutate(voltEXT = ifelse(datetime == "2019-09-13 13:00:00" , mean_voltEXT_4h, voltEXT))

mean_TseaF_4h <- d_all%>%
  dplyr::filter(datetime >= "2019-09-13 11:00:00" & datetime <= "2019-09-13 12:00:00")%>%
  dplyr::summarise( mean(T_seaF, na.rm = TRUE))
mean_TseaF_4h <- as.numeric(mean_TseaF_4h)
d_all <- d_all%>%
  dplyr::mutate(temp_insitu = ifelse(datetime == "2019-09-13 13:00:00" , mean_TseaF_4h, temp_insitu))

#interpolation of salinity during the FB off maintenance
sal_fbinterp <-approx(d_all$datetime, d_all$sal_fb, xout=d_all$datetime, method="linear", rule=2)
names(sal_fbinterp) <- c("datetime", "sal_fb_interp")
sal_fbinterp <- as.data.frame(sal_fbinterp)
d_all$sal_fb_interp <- sal_fbinterp$sal_fb_interp

@@

# test <- d_all%>%
#   dplyr::filter(datetime=="2017-08-24 14:00:00")

@@

# #@@@@@ test plot #@@@@@ #@@@@@ #@@@@@ 
# seafet_contros_xts <- dplyr::select(d_all, datetime, voltINT, phINT_filtered,pH_s_seafet_temp_field, T_seaF, temp_fb, temp_insitu, sal_fb,pHint_tot_sf)
# #seafet_contros_xts <- dplyr::select(d_all, datetime,voltINT)
# seafet_contros_xts <- as.xts(seafet_contros_xts, order.by = d_all$datetime)
# dygraph(seafet_contros_xts, group = "awipev", main="pH seaFET (in situ)", ylab="pHT") %>%
#   dySeries("voltINT", label = "voltINTINT", color ="blue", strokeWidth=0, pointSize=2) %>%
#     dySeries("pH_s_seafet_temp_field", label = "pH spectro", axis= "y2", color =  "green", strokeWidth = 0, pointSize = 4 ) %>%
#     dySeries("pHint_tot_sf", label = "corr pH INT", color ="red",axis= "y2", strokeWidth=0, pointSize=2) %>%
#   dySeries("temp_fb", label = "temp_fb", color ="orange",axis= "y2", strokeWidth=0, pointSize=2) %>%
# 
#   dySeries("sal_fb", label = "sal_fb", color ="pink",axis= "y2", strokeWidth=0, pointSize=2) %>%
#   dySeries("temp_insitu", label = "temp_insitu", color ="black",axis= "y2", strokeWidth=0, pointSize=2) %>%
#     dySeries("T_seaF", label = "T_seaF", color ="grey",axis= "y2", strokeWidth=0, pointSize=2) %>%
#   dyLegend(show = "auto")   %>%
#   dyAxis( "y2",valueRange = c(-3, 36))%>%
#   dyHighlight(highlightCircleSize = 8,highlightSeriesBackgroundAlpha = 0.5,hideOnMouseOut = TRUE) %>%
#   dyOptions(drawGrid = TRUE, drawPoints = TRUE, pointSize = c(1,8,1),useDataTimezone = TRUE, digitsAfterDecimal=4) %>%
#   dyRangeSelector(height = 30, dateWindow= NULL)
# #@@@@@ #@@@@@ #@@@@@ #@@@@@ #@@@@@

# Offset between seafet pH and spec pH 
sfoffset <- left_join(sfPeriod[,c(1,2)], d_all%>%dplyr::select(datetime, pHint_tot_sf), by="datetime")
dat <- mutate(sfoffset, diff_pH_sf = round((pH_s_seafet_temp_field - pHint_tot_sf),digits  = 4))
average_offset <- mean(abs(dat$diff_pH_sf), na.rm=TRUE)

@@
